Tropical Andes plant traits

### 2023-08-04 code/L2/L2_1_TropicalAndes_plant_traits.Rmd
add GIFT trait data

# Read in Data
```{r}
BIEN_traits <- read.csv(file.path(data_path,"TropicalAndes_BIEN_plant_traits_forest.csv"))
TRY_traits <- read.csv(file.path(data_path,"TropicalAndes_TRY_plant_traits_forestSpecies.csv"))
```

@@

```{r}
plant_traits <- dplyr::bind_rows(BIEN_traits_stan, TRY_traits_stan, .id = 'source')
```

```{r}
#replace source 1 or 2 with BIEN or TRY
plant_traits['source'][plant_traits['source'] == 1] <- "BIEN"
plant_traits['source'][plant_traits['source'] == 2] <- "TRY"
```

@@

# Dispersal Traits
```{r}
# subset plant traits to just dispersal traits
plant_traits_dispersal <- plant_traits[plant_traits$trait_name == "Dispersal syndrome" | plant_traits$trait_name == "whole plant dispersal syndrome",]
# look at possible trait values
plant_traits_dispersal_values <- unique(plant_traits_dispersal$trait_value)
# save list to manually classify biotic, abiotic, none, and NA
write.csv(plant_traits_dispersal_values, file.path(data_path,"plant_trait_dispersal_values.csv"))
# read in classified table
dispersal_classified <- read.csv(file.path(data_path,"plant_trait_dispersal_values_classified.csv"))
# reclassify dispersal trait value to table
plant_traits_dispersal$trait_value <- 
dispersal_classified$Classification[match(plant_traits_dispersal$trait_value,dispersal_classified$OrigValue )]
# check to make sure it worked
unique(plant_traits_dispersal$trait_value)
```

@@

# Plant height traits
```{r}
plant_traits_height <- plant_traits[plant_traits$trait_name == "Plant height vegetative" |plant_traits$trait_name == "whole plant height" ,]
plant_traits_height$trait_value <- as.numeric(plant_traits_height$trait_value)
plant_traits_height_wide <- pivot_wider(data = plant_traits_height, id_cols = species, names_from = trait_name, values_from = trait_value, values_fn = mean)
plant_traits_height_wide$plant_height <- rowMeans(plant_traits_height_wide[,2:3], na.rm = TRUE)
# ready to combine
```

# Plant longevity traits
```{r}
plant_traits_longevity <- plant_traits[plant_traits$trait_name == "Plant lifespan (longevity)" | plant_traits$trait_name == "maximum whole plant longevity",]
# change all trait_name to plant lifespan
plant_traits_longevity['trait_name'][plant_traits_longevity['trait_name'] == 'Plant lifespan (longevity)'] <- 'plant_lifespan'
plant_traits_longevity['trait_name'][plant_traits_longevity['trait_name'] == 'maximum whole plant longevity'] <- 'plant_lifespan'

@@

```{r}
plant_traits_wide <- plant_traits_wide [, c("species", "plant_height", "dispersal_syndrome", "Fruit dry mass", "plant_lifespan")]
# remove species with NAs for one or more traits
plant_traits_wide_complete <- na.omit(plant_traits_wide)
```

>>>>

# Read in Data
```{r}
BIEN_traits <- read.csv(file.path(data_path,"TropicalAndes_BIEN_plant_traits_forest.csv"))
TRY_traits <- read.csv(file.path(data_path,"TropicalAndes_TRY_plant_traits_forestSpecies.csv"))
GIFT_traits <- read.csv(file.path(data_path,"TropicalAndes_GIFT_plant_traits_forest.csv"))
```

@@

```{r}
plant_traits <- dplyr::bind_rows(BIEN_traits_stan, TRY_traits_stan, GIFT_traits, .id = 'source')
```

```{r}
#replace source 1 or 2 with BIEN or TRY
plant_traits['source'][plant_traits['source'] == 1] <- "BIEN"
plant_traits['source'][plant_traits['source'] == 2] <- "TRY"
plant_traits['source'][plant_traits['source'] == 3] <- "GIFT"
```

@@

# Dispersal Traits
```{r}
# subset plant traits to just dispersal traits
plant_traits_dispersal <- plant_traits[plant_traits$trait_name == "Dispersal syndrome" | plant_traits$trait_name == "whole plant dispersal syndrome" | plant_traits$trait_name == "dispersal_syndrome",]
# look at possible trait values
plant_traits_dispersal_values <- unique(plant_traits_dispersal$trait_value)
# save list to manually classify biotic, abiotic, none, and NA
write.csv(plant_traits_dispersal_values, file.path(data_path,"plant_trait_dispersal_values.csv"))
plant_traits_dispersal %>%
  count(trait_value)
```

```{r}
# read in classified table
dispersal_classified <- read.csv(file.path(data_path,"plant_trait_dispersal_values_classified.csv"))
# reclassify dispersal trait value to table
plant_traits_dispersal$trait_value <- 
dispersal_classified$Classification[match(plant_traits_dispersal$trait_value,dispersal_classified$OrigValue )]
# check to make sure it worked
unique(plant_traits_dispersal$trait_value)
```

@@

# Plant height traits
```{r}
plant_traits_height <- plant_traits[plant_traits$trait_name == "Plant height vegetative" |plant_traits$trait_name == "whole plant height" |plant_traits$trait_name == "plant_height" ,]
plant_traits_height$trait_value <- as.numeric(plant_traits_height$trait_value)
plant_traits_height_wide <- pivot_wider(data = plant_traits_height, id_cols = species, names_from = trait_name, values_from = trait_value, values_fn = mean)
plant_traits_height_wide$plant_height <- rowMeans(plant_traits_height_wide[,2:3], na.rm = TRUE)
# ready to combine
```

# Plant longevity traits
```{r}
plant_traits_longevity <- plant_traits[plant_traits$trait_name == "Plant lifespan (longevity)" | plant_traits$trait_name == "maximum whole plant longevity" | plant_traits$trait_name == "plant_lifespan",]
# change all trait_name to plant lifespan
plant_traits_longevity['trait_name'][plant_traits_longevity['trait_name'] == 'Plant lifespan (longevity)'] <- 'plant_lifespan'
plant_traits_longevity['trait_name'][plant_traits_longevity['trait_name'] == 'maximum whole plant longevity'] <- 'plant_lifespan'

@@

```{r}
plant_traits_wide <- plant_traits_wide [, c("species", "plant_height", "dispersal_syndrome", "plant_lifespan")]
# remove species with NAs for one or more traits
plant_traits_wide_complete <- na.omit(plant_traits_wide)
```
